Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-10-09 prop_letter.tex
prop_letter.tex \section{Born} Updating...

Recall in Sec.~\ref{sec:fuzzy}, we illustrate the non-constructive
postulates for the quantum probability measures first, and then introduce
the Born rule which induces a quantum probability measures from each
state. Since the postulates for QIVPMs in Sec.~\ref{sec:Interval-Uncertainty}
are also non-constructive, we also want to find a simple rule to construct
a QIVPM. When we generalize quantum probability measures to QIVPMs,
we replace the unit interval range~$\left[0,1\right]$ by a set of
intervals~$\mathscr{I}$. This replacement suggests a function~$\iota$
mapping a number~$x\in\left[0,1\right]$ to an interval~$\iota\left(x\right)\in\mathscr{I}$.
The function~$\iota$ can then be concatenated with a quantum probability
measure~$\mu:\events\rightarrow\left[0,1\right]$ and easily get
an interval-valued function~$\iota\circ\mu:\events\overset{\mu}{\rightarrow}\left[0,1\right]\overset{\iota}{\rightarrow}\mathscr{I}$,
where $\circ$ is the function concatenation. However, the concatenated
function~$\iota\circ\mu$ may or may not be a QIVPM. For example,
in a four-dimensional Hilbert space, consider the function 
\begin{equation}
\iota'\left(x\right)=\begin{cases}
\imposs\,, & \textrm{ if }x=0\,;\\
\left[0,\tfrac{1}{2}\right]\,, & \textrm{ if }x\in\left(0,\tfrac{1}{4}\right]\,;\\
\unknown\,, & \textrm{ if }x\in\left(\tfrac{1}{4},\tfrac{3}{4}\right)\,;\\
\left[\tfrac{1}{2},1\right]\,, & \textrm{ if }x\in\left[\tfrac{3}{4},1\right)\,;\\
\necess\,, & \textrm{ if }x=1\,.
\end{cases}\label{eq:iota}
\end{equation}
While the function~$\iota'\circ\mu_{\rho_{0}}^{\mathrm{\mathrm{B}}}$
induced by $\rho_{0}=\frac{1}{4}\cdot\proj{0}+\frac{1}{4}\cdot\proj{1}+\frac{1}{4}\cdot\proj{2}+\frac{1}{4}\cdot\proj{3}$
is a QIVPM, the function~$\iota'\circ\mu_{\rho_{1}}^{\mathrm{\mathrm{B}}}$
induced by $\rho_{1}=\frac{1}{8}\cdot\proj{0}+\frac{1}{8}\cdot\proj{1}+\frac{1}{8}\cdot\proj{2}+\frac{5}{8}\cdot\proj{3}$
is \emph{not} a QIVPM. Although we could verify whether $\iota\circ\mu$
is a QIVPM or not for any pair of $\iota$ and $\mu$ individually,
the verification process is long and tedious. Therefore, it is desirable
to have some easy criteria to get a QIVPM~$\iota\circ\mu$.

@@

Suppose $\iota$ doesn't satisfy Eq.~(\ref{eq:iota-Inclusion}),
we can find three particular numbers~$x_{0}'$, $x_{1}'$, and $x_{2}'\in\left[0,1\right]$
such that $x_{0}'+x_{1}'+x_{2}'\in\left[0,1\right]$, but 
\[
\iota\left(x_{0}'+x_{1}'+x_{2}'\right)+\iota\left(x_{2}'\right)\nsubseteq\iota\left(x_{0}'+x_{2}'\right)+\iota\left(x_{1}'+x_{2}'\right)\,.
\]
Consider the state
\begin{equation}
\begin{aligned}\rho={} & x_{0}'\proj{0}+x_{1}'\proj{1}+x_{2}'\proj{2}\\
 & +\left(1-x_{0}'-x_{1}'-x_{2}'\right)\proj{3}\,.
\end{aligned}
\end{equation}
Then, the function~$\iota\circ\mu_{\rho}^{\mathrm{\mathrm{B}}}:\events\rightarrow\mathscr{I}$
fails to satisfy Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
\[
\begin{aligned} & \mathrel{\hphantom{\nsubseteq}}\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{1}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{2}\right)\right)\\
 & \nsubseteq\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{1}+\proj{2}\right)\right)\,,
\end{aligned}
\]
and cannot be a QIVPM. Therefore, to make sure $\iota\circ\mu$ is
always a QIVPM, $\iota:\left[0,1\right]\rightarrow\mathscr{I}$ should
satisfy Eqs.~(\ref{eq:iota-constraints}) and Eq.~(\ref{eq:iota-Inclusion}).

Since $\iota'\circ\mu_{\rho_{1}}^{\mathrm{\mathrm{B}}}$ is not a
QIVPM, we know $\iota'$ doesn't satisfy Eq.~(\ref{eq:iota-Inclusion}).
In order to find $\iota$ satisfying Eq.~(\ref{eq:iota-Inclusion}),
we decompose it into two functions: its left-end and right-end~$\left[\iotal\left(x\right),\iotar\left(x\right)\right]=\iota\left(x\right)$.
They are both usual functions from a unit interval to another unit
interval. Eq.~(\ref{eq:iota-Inclusion}) then implies that $\iotal$
is Wright-convex, i.e., $\iotal\left(x_{0}+x_{1}+x_{2}\right)+\iotal\left(x_{2}\right)\ge\iotal\left(x_{0}+x_{2}\right)+\iotal\left(x_{1}+x_{2}\right)$,
and $\iotar$ is Wright-concave, i.e., $\iotar\left(x_{0}+x_{1}+x_{2}\right)+\iotar\left(x_{2}\right)\le\iotar\left(x_{0}+x_{2}\right)+\iotar\left(x_{1}+x_{2}\right)$~\citep{Wright1954,RobertsVarberg1973,PecaricTong1992,Brzdek_2014}.
Notice that a Wright-convex function is midpoint-convex or Jensen-convex,
i.e., 
\begin{equation}
\frac{\iotal\left(x_{0}\right)+\iotal\left(x_{1}\right)}{2}\ge\iotal\left(\frac{x_{0}+x_{1}}{2}\right)\,,
\end{equation}
and a midpoint-convex function is continuous on an open interval if
we cannot use the Axiom of Choices~\citep{Blumberg_1919,Sierpinski1920,RobertsVarberg1973}.
Therefore, the range of $\iota$ either is $\left\{ \necess,\imposs,\unknown\right\} $
or contains infinitely many intervals.

>>>>

Recall in Sec.~\ref{sec:fuzzy}, we illustrate the non-constructive
postulates for the quantum probability measures first, and then introduce
the Born rule which induces a quantum probability measures from each
state. Since the postulates for QIVPMs in Sec.~\ref{sec:Interval-Uncertainty}
are also non-constructive, we want to find a simple rule to construct
a QIVPM as well. When we generalized quantum probability measures
to QIVPMs, we changed the range from the unit interval~$\left[0,1\right]$
to a set of intervals~$\mathscr{I}$. This replacement suggests us
to find a function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$
which can then be concatenated with a quantum probability measure~$\mu:\events\rightarrow\left[0,1\right]$
and easily produce an interval-valued function~$\iota\circ\mu:\events\overset{\mu}{\rightarrow}\left[0,1\right]\overset{\iota}{\rightarrow}\mathscr{I}$,
where $\circ$ is the function concatenation. However, the concatenated
function~$\iota\circ\mu$ may or may not be a QIVPM. For example,
in a four-dimensional Hilbert space, consider the function 
\begin{equation}
\iota'\left(x\right)=\begin{cases}
\imposs\,, & \textrm{ if }x=0\,;\\
\left[0,\tfrac{1}{2}\right]\,, & \textrm{ if }x\in\left(0,\tfrac{1}{4}\right]\,;\\
\unknown\,, & \textrm{ if }x\in\left(\tfrac{1}{4},\tfrac{3}{4}\right)\,;\\
\left[\tfrac{1}{2},1\right]\,, & \textrm{ if }x\in\left[\tfrac{3}{4},1\right)\,;\\
\necess\,, & \textrm{ if }x=1\,.
\end{cases}\label{eq:iota}
\end{equation}
While the function~$\iota'\circ\mu_{\rho_{0}}^{\mathrm{\mathrm{B}}}$
induced by $\rho_{0}=\frac{1}{4}\cdot\proj{0}+\frac{1}{4}\cdot\proj{1}+\frac{1}{4}\cdot\proj{2}+\frac{1}{4}\cdot\proj{3}$
is a QIVPM, the function~$\iota'\circ\mu_{\rho_{1}}^{\mathrm{\mathrm{B}}}$
induced by $\rho_{1}=\frac{1}{8}\cdot\proj{0}+\frac{1}{8}\cdot\proj{1}+\frac{1}{8}\cdot\proj{2}+\frac{5}{8}\cdot\proj{3}$
is \emph{not} a QIVPM. Although we could verify whether $\iota\circ\mu$
is a QIVPM or not for any pair of $\iota$ and $\mu$ individually,
the verification process is long and tedious. Therefore, it is desirable
to have some easy criteria to decide when $\iota\circ\mu$ is a QIVPM.

@@

The proof is by contradiction: Suppose $\iota$ doesn't satisfy Eq.~(\ref{eq:iota-Inclusion}),
we can find three particular numbers~$x_{0}$, $x_{1}$, and $x_{2}\in\left[0,1\right]$
such that $x_{0}+x_{1}+x_{2}\in\left[0,1\right]$, but 
\[
\iota\left(x_{0}+x_{1}+x_{2}\right)+\iota\left(x_{2}\right)\nsubseteq\iota\left(x_{0}+x_{2}\right)+\iota\left(x_{1}+x_{2}\right)\,.
\]
Consider the state $\rho=x_{0}\proj{0}+x_{1}\proj{1}+x_{2}\proj{2}+\left(1-x_{0}-x_{1}-x_{2}\right)\proj{3}$.
Then, the function~$\iota\circ\mu_{\rho}^{\mathrm{\mathrm{B}}}:\events\rightarrow\mathscr{I}$
fails to satisfy Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
\begin{eqnarray*}
 &  & \iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{1}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{2}\right)\right)\\
 & \nsubseteq & \iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{1}+\proj{2}\right)\right)\,,
\end{eqnarray*}
and cannot be a QIVPM. Therefore, to make sure $\iota\circ\mu$ is
always a QIVPM, $\iota:\left[0,1\right]\rightarrow\mathscr{I}$ should
satisfy Eqs.~(\ref{eq:iota-constraints}) and Eq.~(\ref{eq:iota-Inclusion}).

Since $\iota'\circ\mu_{\rho_{1}}^{\mathrm{\mathrm{B}}}$ is not a
QIVPM, we know $\iota'$ doesn't satisfy Eq.~(\ref{eq:iota-Inclusion}).
In order to find a function~$\iota$ satisfying Eq.~(\ref{eq:iota-Inclusion}),
we decompose it into two functions: its left-end and right-end~$\left[\iotal\left(x\right),\iotar\left(x\right)\right]=\iota\left(x\right)$.
They are both usual functions from a unit interval to another unit
interval. Eq.~(\ref{eq:iota-Inclusion}) then implies that $\iotal$
is Wright-convex, i.e., $\iotal\left(x_{0}+x_{1}+x_{2}\right)+\iotal\left(x_{2}\right)\ge\iotal\left(x_{0}+x_{2}\right)+\iotal\left(x_{1}+x_{2}\right)$,
and $\iotar$ is Wright-concave, i.e., $\iotar\left(x_{0}+x_{1}+x_{2}\right)+\iotar\left(x_{2}\right)\le\iotar\left(x_{0}+x_{2}\right)+\iotar\left(x_{1}+x_{2}\right)$~\citep{Wright1954,RobertsVarberg1973,PecaricTong1992,Brzdek_2014}.
A Wright-convex function is midpoint-convex or Jensen-convex~\citep{RobertsVarberg1973,PecaricTong1992},
i.e., we have
\begin{equation}
\frac{\iotal\left(x_{0}\right)+\iotal\left(x_{1}\right)}{2}\ge\iotal\left(\frac{x_{0}+x_{1}}{2}\right)\,.
\end{equation}
A midpoint-convex function is continuous on an open interval if we
cannot use the Axiom of Choices~\citep{Blumberg_1919,Sierpinski1920,RobertsVarberg1973}.
Therefore, the range of $\iota$ either is $\left\{ \necess,\imposs,\unknown\right\} $
or contains infinitely many intervals.

\yutsung{Another idea might be fixing a quantum probability measure~$\mu$,
whether there is a length of intervals, such that if $\iota$ mapping
to a interval tighter than the given length, then $\iota\circ\mu$
is always a QIVPM... I don't know the answer of this question.}
